---
title: "Fantastic Four"
author: "Rosa Choe, Adam Dada, Jen Shi, Katherine Sheng"
date: "May 2, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(png)
library(dplyr)
library(ggplot2)
```
## Research Topic

- Questions: How does UC Berkeley compare to other UCs in terms of safety? How do NorCal schools differ from SoCal public 4-yr universities?
- Variables of Interest: Location (on campus, off campus, student housing facilities), Categories of Crime (arrests, disciplinary actions, criminal offenses), Region (NorCal vs. SoCal)
- Data Source: U.S. Dept of Education [https://ope.ed.gov/campussafety/#/customdata/search], Gallup [http://www.gallup.com/poll/194693/parents-fears-child-safety-school-unchanged.aspx]
- Methods: Data Wrangling (regex, filter), Data Visualization (ggplot), 'for' loop, Defined Functions

## Berkeley vs. UCs

```{r}
UC_arrests_noncampus = read.csv("NorCal SoCal/Public_UC_Arrests_Noncampus.csv")
UC_arrests_campus = read.csv("NorCal SoCal/Public_UC_Arrests_On_Campus.csv")
UC_arrests_housing = read.csv("NorCal SoCal/Public_UC_Arrests_On_campus_Student_Housing_Facilities.csv")

UC_criminal_noncampus = read.csv("NorCal SoCal/Public_UC_Criminal_Offenses_Noncampus.csv")
UC_criminal_campus = read.csv("NorCal SoCal/Public_UC_Criminal_Offenses_On_campus.csv")
UC_criminal_housing = read.csv("NorCal SoCal/Public_UC_Criminal_Offenses_On_campus_Student_Housing_Facilities.csv")

UC_disciplinary_noncampus = read.csv("NorCal SoCal/Public_UC_Disciplinary_Actions_Noncampus.csv")
UC_disciplinary_campus = read.csv("NorCal SoCal/Public_UC_Disciplinary_Actions_On_campus.csv")
UC_disciplinary_housing = read.csv("NorCal SoCal/Public_UC_Disciplinary_Actions_Student_Housing_Facilities.csv")

NUC_arrests_noncampus = read.csv("NorCal SoCal/Public_NUC_Arrests_Noncampus.csv")
NUC_arrests_campus = read.csv("NorCal SoCal/Public_NUC_Arrests_On_campus.csv")
NUC_arrests_housing = read.csv("NorCal SoCal/Public_NUC_Arrests_On_campus_Student_Housing_Facilities.csv")

NUC_criminal_noncampus = read.csv("NorCal SoCal/Public_NUC_Criminal_Offenses_Noncampus.csv")
NUC_criminal_campus = read.csv("NorCal SoCal/Public_NUC_Criminal_Offenses_On_campus.csv")
NUC_criminal_housing = read.csv("NorCal SoCal/Public_NUC_Criminal_Offenses_On_campus_Student_Housing_Facilities.csv")

NUC_disciplinary_noncampus = read.csv("NorCal SoCal/Public_NUC_Disciplinary_Actions_Noncampus.csv")
NUC_disciplinary_campus = read.csv("NorCal SoCal/Public_NUC_Disciplinary_Actions_On_campus.csv")
NUC_disciplinary_housing = read.csv("NorCal SoCal/Public_NUC_Disciplinary_Actions_Student_Housing_Facilities.csv")
Arrests <- c("Weapons", "Drugs", "Liquor")
ArrestsPerStudent <- paste(Arrests, "PerStudent", sep="")

Crimes <- c("ForcibleSexOffenses", "Assault", "Burglary", "CarTheft", "Arson", "Total")
CrimesPerStudent <- paste(Crimes, "PerStudent", sep="")

VAWA <- c("DomesticViolence", "DatingViolence", "Stalking", "Rape", "Fondling", "StatutoryRape")
VAWAPerStudent <- paste(VAWA, "PerStudent", sep="")

filterArrests <- function(data) {
  return(data %>% filter(!grepl("Hastings", data$Institution.name)) %>%
                                          filter(!is.na(Institution.Size)) %>% 
                                          mutate(Weapons = as.numeric(Illegal.weapons.possession),
                                                 Drugs = as.numeric(Drug.law.violations),
                                                 Liquor = as.numeric(Liquor.law.violations)) %>%
                                          mutate(WeaponsPerStudent = Weapons/Institution.Size, 
                                                 DrugsPerStudent = Drugs/Institution.Size, 
                                                 LiquorPerStudent = Liquor/Institution.Size) %>%
                                          mutate(Total = Weapons + Drugs + Liquor, TotalPerStudent = Total/Institution.Size) %>%
                                          select(Survey.year, Institution.name, one_of(Arrests), one_of(ArrestsPerStudent)))
}

filterCriminal <- function(data){
   data %>% filter(!grepl("Hastings", data$Institution.name)) %>%
                                            filter(!is.na(Institution.Size)) %>% 
                                            mutate(ForcibleSexOffenses = as.numeric(Sex.offenses...Forcible),
                                                   Assault = as.numeric(Aggravated.assault),
                                                   CarTheft = as.numeric(Motor.vehicle.theft)) %>%
                                            mutate(ForcibleSexOffensesPerStudent = ForcibleSexOffenses/Institution.Size,
                                                   AssaultPerStudent = Assault/Institution.Size,
                                                   BurglaryPerStudent = Burglary/Institution.Size,
                                                   CarTheftPerStudent = CarTheft/Institution.Size,
                                                   ArsonPerStudent = Arson/Institution.Size) %>%
                                            mutate(Total = ForcibleSexOffenses + Assault + Burglary + CarTheft + Arson,
                                                   TotalPerStudent = Total/Institution.Size) %>%
                                            select(Survey.year, Institution.name, one_of(Crimes), one_of(CrimesPerStudent))
}

filterVAWA <- function(data) {
  data %>% filter(!grepl("Hastings", data$Institution.name)) %>%
                                            filter(!is.na(Institution.Size)) %>% 
                                            mutate(DomesticViolence = as.numeric(Domestic.violence),
                                                   DatingViolence = as.numeric(Dating.violence),
                                                   StatutoryRape = as.numeric(Statutory.Rape)) %>%
                                            mutate(DomesticViolencePerStudent = DomesticViolence/Institution.Size,
                                                   DatingViolencePerStudent = DatingViolence/Institution.Size,
                                                   StalkingPerStudent = Stalking/Institution.Size,
                                                   RapePerStudent = Rape/Institution.Size,
                                                   FondlingPerStudent = Fondling/Institution.Size,
                                                   StatutoryRapePerStudent = StatutoryRape/Institution.Size) %>%
                                            mutate(Total = DomesticViolence + DatingViolence + Stalking + Rape + Fondling + StatutoryRape) %>%
                                            select(Survey.year, Institution.name, one_of(VAWA), one_of(VAWAPerStudent))
}

sumSexOffenses <- function(data) {
  return(data %>% mutate(Sex.offenses...Forcible = ifelse(Survey.year > 2013, Rape + Fondling, Sex.offenses...Forcible)) %>%
          mutate(Sex.offenses...Non.forcible = ifelse(Survey.year > 2013, Incest + Statutory.rape, Sex.offenses...Non.forcible)))
}

updatedNames <- function(data, UC=TRUE) {
  if (UC) {
    return(gsub(".*-", "UC ", data$Institution.name))
  } else {
    data$Institution.name <- gsub("California State University", "CSU", data$Institution.name)
    data$Institution.name <- gsub(".*Polytechnic.*-", "CalPoly-", data$Institution.name)
    return(data$Institution.name)
  }
  
}

updateNames <- function(data, UC=TRUE) {
  if (UC) {
    data$name = gsub(".*-", "", data$name)
  } else {
    data$name <- gsub("California State University", "CSU", data$name)
    data$name <- gsub(".*Polytechnic.*-", "CalPoly-", data$name)
  }
  return(data)
  
}

getBerkeley <- function(data) {
  return(data %>% filter(grepl("Berkeley", Institution.name)) %>% ungroup())
}

sumNonBerkeley <- function(data, UC=TRUE) {
  data = data %>% filter(!grepl("Berkeley", Institution.name)) %>% 
                                                group_by(Survey.year) %>% 
                                                summarise_each(funs(mean))

  if (UC) {
     return(data %>% mutate(Institution.name = "Other UCs"))
  } else {
    return(data %>% mutate(Institution.name = "Non-UCs"))
  }
}

socalNorcal <- function(data1, data2, group1, group2) {
  data1 = data1 %>% na.omit() %>%
                  group_by(Survey.year) %>% 
                  summarise_each(funs(mean)) %>%
                  mutate(Institution.name = group1)
  
  data2 = data2 %>% group_by(Survey.year) %>% 
                  na.omit() %>%
                  summarise_each(funs(mean)) %>%
                  mutate(Institution.name = group2)

  return(rbind(data1, data2))
}

massPlot <- function(data, cols) {
  for(col in cols) {
    print(data %>% ggplot(aes(x = Survey.year, col=Institution.name)) + geom_line(aes_string(y=col)))
  }
}

massPlot2 <- function(data, cols) {
  for(col in cols) {
    print(data %>% ggplot(aes(x = year, col=name)) + geom_line(aes_string(y=col)))
  }
}

groupCriminal <- function(data1, data2, data3, UC=TRUE){
  data <- rbind(data1, data2, data3)
  data$Institution.name = updatedNames(data, UC)
  return(data %>% sumSexOffenses() %>%
                                filterCriminal() %>%
                                group_by(Survey.year, Institution.name) %>%
                                summarise_all(.funs=sum) %>%
                                ungroup(Survey.year, Institution.name))
}

groupArrests <- function(data1, data2, data3, UC=TRUE){
  data <- rbind(data1, data2, data3)
  data$Institution.name = updatedNames(data, UC)
  return(data %>% filterArrests() %>%
                                group_by(Survey.year, Institution.name) %>%
                                summarise_all(.funs=sum) %>%
                                ungroup(Survey.year, Institution.name))
}
UC_criminal <- groupCriminal(UC_criminal_campus, UC_criminal_housing, UC_criminal_noncampus, UC=TRUE)

UC_arrests <- groupArrests(UC_arrests_campus, UC_arrests_housing, UC_arrests_noncampus, UC=TRUE)

UC_disciplinary <- groupArrests(UC_disciplinary_campus, UC_disciplinary_housing, UC_disciplinary_noncampus, UC=TRUE)

NUC_criminal <- groupCriminal(NUC_criminal_campus, NUC_criminal_housing, NUC_criminal_noncampus, UC=FALSE)

NUC_arrests <- groupArrests(NUC_arrests_campus, NUC_arrests_housing, NUC_arrests_noncampus, UC=FALSE)

NUC_disciplinary <- groupArrests(NUC_disciplinary_campus, NUC_disciplinary_housing, NUC_disciplinary_noncampus, UC=FALSE)
Berkeley_arrests = getBerkeley(UC_arrests)

Other_UC_arrests = sumNonBerkeley(UC_arrests)

summedUC_arrests = rbind(Berkeley_arrests, Other_UC_arrests)

massPlot(UC_arrests, Arrests)
massPlot(UC_arrests, ArrestsPerStudent)

massPlot(summedUC_arrests, Arrests)
massPlot(summedUC_arrests, ArrestsPerStudent)
Berkeley_criminal = getBerkeley(UC_criminal)

Other_UC_criminal = sumNonBerkeley(UC_criminal)

summedUC_criminal = rbind(Berkeley_criminal, Other_UC_criminal)

massPlot(UC_criminal, Crimes)
massPlot(UC_criminal, CrimesPerStudent)

massPlot(summedUC_criminal, Crimes)
massPlot(summedUC_criminal, CrimesPerStudent)
Berkeley_disciplinary = getBerkeley(UC_disciplinary)

Other_UC_disciplinary = sumNonBerkeley(UC_disciplinary)

summedUC_disciplinary = rbind(Berkeley_disciplinary, Other_UC_disciplinary)

massPlot(UC_disciplinary, Arrests)
massPlot(UC_disciplinary, ArrestsPerStudent)

massPlot(summedUC_disciplinary, Arrests)
massPlot(summedUC_disciplinary, ArrestsPerStudent)
compareDiscplinaryArrests <- function(disciplinary, arrests) {
  combined <- merge(disciplinary, arrests, by=c("Survey.year", "Institution.name"))
  UC_criminal$Institution.name = gsub(".*-", "", UC_criminal$Institution.name)
  names(combined) <- gsub("y$","Arrests", names(combined))
  names(combined) <- gsub("x$","Disciplinary", names(combined))
  for(crime in Arrests) {
    crime_x = paste(crime, "Disciplinary", sep=".")
    crime_y = paste(crime, "Arrests", sep=".")
    print(combined %>% ggplot(aes(col=Institution.name)) + geom_point(aes_string(x=crime_x, y=crime_y)))
  }
}

compareDiscplinaryArrests(UC_disciplinary, UC_arrests)
```

## NorCal vs. SoCal

```{r}
Berkeley_NUC_arrests = rbind(Berkeley_arrests, NUC_arrests)


Berkeley_summedNUC_arrests = rbind(Berkeley_arrests, sumNonBerkeley(NUC_arrests, FALSE))

massPlot(NUC_arrests, Arrests)
massPlot(NUC_arrests, ArrestsPerStudent)

massPlot(Berkeley_NUC_arrests, Arrests)
massPlot(Berkeley_NUC_arrests, ArrestsPerStudent)

massPlot(Berkeley_summedNUC_arrests, Arrests)
massPlot(Berkeley_summedNUC_arrests, ArrestsPerStudent)
Berkeley_NUC_criminal = rbind(Berkeley_criminal, NUC_criminal)

Berkeley_summedNUC_criminal = rbind(Berkeley_criminal, sumNonBerkeley(NUC_criminal, FALSE))


massPlot(Berkeley_NUC_criminal, Crimes)
massPlot(Berkeley_NUC_criminal, CrimesPerStudent)

massPlot(Berkeley_summedNUC_criminal, Crimes)
massPlot(Berkeley_summedNUC_criminal, CrimesPerStudent)
Berkeley_NUC_disciplinary = rbind(Berkeley_disciplinary, NUC_disciplinary)

Berkeley_summedNUC_disciplinary = rbind(Berkeley_disciplinary, sumNonBerkeley(NUC_disciplinary, FALSE))

massPlot(Berkeley_NUC_disciplinary, Arrests)
massPlot(Berkeley_NUC_disciplinary, ArrestsPerStudent)

massPlot(Berkeley_summedNUC_disciplinary, Arrests)
massPlot(Berkeley_summedNUC_disciplinary, ArrestsPerStudent)
NorCal_UC = c("UC Berkeley", "UC Davis", "UC San Francisco", "UC Santa Cruz", "UC Merced")

#NorCal UC datatables
NorCal_UC_arrests <- UC_arrests %>%
  filter(Institution.name %in% NorCal_UC)

NorCal_UC_disciplinary <- UC_disciplinary %>%
  filter(Institution.name %in% NorCal_UC)

NorCal_UC_criminal <- UC_criminal %>%
  filter(Institution.name %in% NorCal_UC)

#SoCal UC datatables
SoCal_UC_arrests <- UC_arrests %>%
  filter(!Institution.name %in% NorCal_UC)

SoCal_UC_disciplinary <- UC_disciplinary %>%
  filter(!Institution.name %in% NorCal_UC)

SoCal_UC_criminal <- UC_criminal %>%
  filter(!Institution.name %in% NorCal_UC)
NorCal_NUC = c("CSU-Chico", "CSU-Sacramento", "San Jose State University", "Sonoma State University", "Humboldt State University")

#NorCal CSU datatables
NorCal_NUC_arrests <- NUC_arrests %>%
        filter(Institution.name %in% NorCal_NUC)

NorCal_NUC_disciplinary <- NUC_disciplinary %>%
        filter(Institution.name %in% NorCal_NUC)
  
NorCal_NUC_criminal <- NUC_criminal %>%
        filter(Institution.name %in% NorCal_NUC)

#SoCal UC datatables
SoCal_NUC_arrests <- NUC_arrests %>%
        filter(!Institution.name %in% NorCal_NUC)

SoCal_NUC_disciplinary <- NUC_disciplinary %>%
        filter(!Institution.name %in% NorCal_NUC)

SoCal_NUC_criminal <- NUC_criminal %>%
        filter(!Institution.name %in% NorCal_NUC)
NorcalSocal_UC_criminal = socalNorcal(NorCal_UC_criminal, SoCal_UC_criminal, "Norcal", "Socal")

massPlot(NorcalSocal_UC_criminal, Crimes)
NorcalSocal_UC_arrests = socalNorcal(NorCal_UC_arrests, SoCal_UC_arrests, "Norcal", "Socal")

massPlot(NorcalSocal_UC_arrests, Arrests)
NorcalSocal_UC_disciplinary = socalNorcal(NorCal_UC_disciplinary, SoCal_UC_disciplinary, "Norcal", "Socal")

massPlot(NorcalSocal_UC_disciplinary, Arrests)
```

